### 2021-12-13 documentos/generators.tex
Algunas frases sobre simetría de PCEGs

%\colorbox{yellow}{\textcolor{blue}{Esqueleto}}
%\textcolor{jacolor}{
%\begin{itemize}
%\item Siguiendo la discusión de los espacios vectoriales de la sección anterior, 
%introducir de forma consecuente los generadores como subespacios maximales. 
%Es decir, darle una retocada a lo que ya escribió Francois, pero que armonice 
%con lo que viene de la sección anterior.
%\item Ilustrar el algoritmo para construir todos los PCEs como 
%concatenaciones de los generadores. Comentar al final del párrafo que, para 
%1 qubit, nuestro descubrimiento de los PCEGs es compatible con los resultados 
%de Siudzinska~\cite{siudzinska2019geometry} sobre el espacio en el tetrahedro 
%en el que ella encontró que viven los canales cuánticos que son accesibles 
%mediante generadores. **en algún lado aquí se puede decir que en el apéndice 
%se muestra con los tableros como generar todo el caso de 2 qubits
%\item Ya que expusimos en el párrafo anterior la utilidad de los generadores, 
%hablar de la relación de los generadores con la $A$, como una forma de 
%obtenerlos. El enunciado principal podría ser algo como "The matrix $A$ 
%contains all the information of PCE quantum channels". Es una frase fuerte, 
%pero creo que con un poco de meditación es cierto; de la $A$ puedes sacar 
%los generadores, y con ellos todos los PCEs, se acabó. 
%\item Hablar de las simetrías de los generadores. (No sé a qué se refiere quien 
%escribió eso. Toca hablarlo con Carlos.)
%\end{itemize}}

Un párrafo de intro a las tres ideas 
\begin{itemize}
\item Existen los generadores 
\item Clasificación de los generadores 
\item Simetrías
\end{itemize}

@@

\cpnote{Arreglar}
PCE generators may be characterized by its local-action on every qubit 
in the system, and thus be denoted as $\mcG_{\vec\alpha}$. First, let us denote
the four 1-qubit PCE generators as $\mcG_{\alpha}$, with $\alpha=0,1,2,3$.
The first being the trivial generator and the remaining the channels that 
map the Bloch sphere to each of the Cartesian axes.
%PCE generator may be identified with a unique vector index
%$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
%Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
%the identity map and the 3 1-qubit PCE quantum channels that project the 
%Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
%If one sets all indices in $\vec\alpha$ to zero except for the $k$-th index.
\janote{A rigorous explanation of why every $\vec\alpha$ is corresponded
to every generator.}
Therefore we can identify the index $\alpha_k$ in $\vec\alpha$ with the local-action
of the PCE generator on the $k$-th qubit. In consequence, we will denote 
PCE generators as $\mcG_{\vec\alpha}$. 
See in \Fref{fig:appex:twoQ_PCEGs} the notation of  two-qubit PCE generators.
%See, for example, non-trivial generators in subfigures (c), (d) and (e) of
%\Fref{fig:examplesPCE} are denoted $\mcG_{(1,3)}$, $\mcG_{(3,2)}$ and
%$\mcG_{(2,2)}$, respectively.
%whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
%to the $x$ axis and on the second qubit (first row) to map the corresponding
%local Bloch sphere to itself, may be characterized with the vector index (1,0). 
%Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
%every PCE generator and so denote them as $\pceg$.

\cpnote{Mucha lora y no se prueba nada}
The same symmetry can be identified for both the PCE generators $\pceg$ and
matrix $A$ the diagonalizes PCE map's Choi matrix $\mcD_N$.
%PCE generators $\pceg$ obey a symmetry
%that defines uniquely every PCE generator. 
For a PCE generator the symmetry is encoded in the array of $\taus$, and 
for $A$ in its rows and columns.\janote{aquí voy}
The coefficients  $\{\taus\}$ of every 
generator $\pceg$ arranged in a hypercubic $N$-dimensional array 
%of dimensions $\underbrace{4\times 4\times\ldots\times 4}_{N}$ 
obey a symmetry with respect to the $N$ hyperplanes of dimension $N-1$
\janote{según wikipedia, por definición, los hiperplanos son de dimensión 
$N-1$. Así que podría quitarse la aclaración de la dimensión. Lo dejo así y 
espero retroalimentación} 
that separate the array into two half arrays \janote{No estoy tan satisfecho con la redacción de este enunciado}.
See, in \Fref{fig:examplesPCE}, that generators $\mcG_{(1,3)}$, $\mcG_{(3,2)}$
and $\mcG_{(2,2)}$ represented in (c), (d) and (e), respectively,
are either symmetric or anti-symmetric with respect to the lines that cut vertically 
and horizontally in half the diagrams. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows.\janote{sintetizar} Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.

>>>>

%\colorbox{yellow}{\textcolor{blue}{Esqueleto}}
%\textcolor{jacolor}{
%\begin{itemize}
%\item Siguiendo la discusión de los espacios vectoriales de la sección anterior, 
%introducir de forma consecuente los generadores como subespacios maximales. 
%Es decir, darle una retocada a lo que ya escribió Francois, pero que armonice 
%con lo que viene de la sección anterior.
%\item Ilustrar el algoritmo para construir todos los PCEs como 
%concatenaciones de los generadores. Comentar al final del párrafo que, para 
%1 qubit, nuestro descubrimiento de los PCEGs es compatible con los resultados 
%de Siudzinska~\cite{siudzinska2019geometry} sobre el espacio en el tetrahedro 
%en el que ella encontró que viven los canales cuánticos que son accesibles 
%mediante generadores. **en algún lado aquí se puede decir que en el apéndice 
%se muestra con los tableros como generar todo el caso de 2 qubits
%\item Ya que expusimos en el párrafo anterior la utilidad de los generadores, 
%hablar de la relación de los generadores con la $A$, como una forma de 
%obtenerlos. El enunciado principal podría ser algo como "The matrix $A$ 
%contains all the information of PCE quantum channels". Es una frase fuerte, 
%pero creo que con un poco de meditación es cierto; de la $A$ puedes sacar 
%los generadores, y con ellos todos los PCEs, se acabó. 
%\item Hablar de las simetrías de los generadores. (No sé a qué se refiere quien 
%escribió eso. Toca hablarlo con Carlos.)
%\end{itemize}}
%
%Un párrafo de intro a las tres ideas 
%\begin{itemize}
%\item Existen los generadores 
%\item Clasificación de los generadores 
%\item Simetrías
%\end{itemize}

@@

%\cpnote{Arreglar}
\janote{Discutir con Carlos}
PCE generators may be characterized by its local-action on every qubit 
in the system, and thus be denoted as $\mcG_{\vec\alpha}$. First, let us denote
the four 1-qubit PCE generators as $\mcG_{\alpha}$, with $\alpha=0,1,2,3$.
The first being the trivial generator and the remaining the channels that 
map the Bloch sphere to each of the Cartesian axes.
%PCE generator may be identified with a unique vector index
%$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
%Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
%the identity map and the 3 1-qubit PCE quantum channels that project the 
%Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
%If one sets all indices in $\vec\alpha$ to zero except for the $k$-th index.
\janote{A rigorous explanation of why every $\vec\alpha$ is corresponded
to every generator.}
Therefore we can identify the index $\alpha_k$ in $\vec\alpha$ with the local-action
of the PCE generator on the $k$-th qubit. In consequence, we will denote 
PCE generators as $\mcG_{\vec\alpha}$. 
See in \Fref{fig:appex:twoQ_PCEGs} the notation of  two-qubit PCE generators.
%See, for example, non-trivial generators in subfigures (c), (d) and (e) of
%\Fref{fig:examplesPCE} are denoted $\mcG_{(1,3)}$, $\mcG_{(3,2)}$ and
%$\mcG_{(2,2)}$, respectively.
%whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
%to the $x$ axis and on the second qubit (first row) to map the corresponding
%local Bloch sphere to itself, may be characterized with the vector index (1,0). 
%Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
%every PCE generator and so denote them as $\pceg$.

%\cpnote{Mucha lora y no se prueba nada}
A symmetry is identified for both PCE generators $\pceg$ and
the matrix $A$ the diagonalizes a PCE map's Choi matrix $\mcD_N$.
%PCE generators $\pceg$ obey a symmetry
%that defines uniquely every PCE generator. 
For a PCE generator the symmetry is encoded in the parameters $\{\taus\}$, and 
for $A$ in its rows and columns.
The $N$-dimensional Cartesian grid associated with $\{\tau_{\vec\alpha}\}$
is either symmetric or anti-symmetric with respect to
the $N$ hyperplanes of dimension $N-1$ that divide the grid in half.
For instance, see that the diagrams representing the generators 
$\mcG_{(1,3)}$, $\mcG_{(3,2)}$ and $\mcG_{(2,2)}$ in subfigures 
(c), (d) and (e) of \Fref{fig:two:qubit:examples}, respectively,
are either symmetric or anti-symmetric with respect to the lines that divide vertically 
and horizontally the diagrams in half. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows.\janote{sintetizar} Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.
